---
title: "DV_4_homework_CR"
author: "Christoph Rosa"
date: "2022-09-13"
output: html_document
---

# Set-up


Installing packages:
```{r}
library(tidyverse)
```

Downloading and cleaning up data:
```{r}
tvratings <- read_csv("../Data/tv_ratings.csv")
glimpse(tvratings)
```


# Question 1

First, let's filter out the "short" shows:
```{r}
tvlong <- tvratings %>% 
  group_by(title) %>% 
  mutate(num_seasons = max(seasonNumber)) %>% 
  ungroup() %>% 
  filter(num_seasons >= 5)

```

And now we plot their ratings over the course of their run:
```{r}
ggplot(tvlong,
       aes(seasonNumber, av_rating,
           group = title)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Average rating of show by season",
       x = "Season",
       y = "Average rating")
```

Honestly, I can't.


# Question 2

Faceting by genre:
```{r}
ggplot(tvlong,
       aes(seasonNumber, av_rating,
           group = title)) +
  geom_line() +
  theme_minimal() +
    labs(title = "Average rating of show by season",
       x = "Season",
       y = "Average rating") +
  facet_wrap(~ genres)
```

Now it becomes apparent that, in the majority of cases, the ratings of shows declined as the seasons progress. 

```{r}
tvlong %>% 
  filter(genres == "Drama,Family,Fantasy") %>% 
  select(title, av_rating) 
```

Looks like "Are you Afraid of the Dark?" (whatever that is) really lost it in its last two seasons. 


# Question 3

Plotting the best-rated shows:
```{r}
best <- tvratings %>% 
  filter(av_rating >= 9)

ggplot(best,
       aes(genres)) +
  geom_bar() +
  theme_minimal() +
  labs(title = "Number of top-rated shows per genre",
       x = "Genre",
       y = "# of shows with average rating > 9.0") +
  scale_y_continuous(breaks = seq(0, 14, 2))
```

Flipping the barplot:
```{r}
ggplot(best,
       aes(genres)) +
  geom_bar() +
  theme_minimal() +
  labs(title = "Number of top-rated shows per genre",
       x = "Genre",
       y = "# of shows with average rating > 9.0") +
  scale_y_continuous(breaks = seq(0, 14, 2)) +
  coord_flip()
```

The genre drama has the most top-rated shows. Not very surprising since all the other genre-combinations are very specific. 


# Question 4

Coding all comedy-somethings as comedies: 
```{r}
comedies_dramas <- tvratings %>% 
  mutate(is_comedy = if_else(str_detect(genres, "Comedy"), 
                             1, 
                             0)) %>% # If it contains the word comedy then 1, else 0
  filter(is_comedy == 1 | genres == "Drama") %>% # Keep comedies and dramas
  mutate(genres = if_else(genres == "Drama", # Make it so that we only have those two genres
                          "Drama", 
                          "Comedy"))

glimpse(comedies_dramas)
```

Comparing the two genres:
```{r}
ggplot(comedies_dramas,
       aes(av_rating, color = genres)) +
  geom_density() +
  theme_minimal() +
  labs(title = "Distribution of average ratings by genre",
       x = "Average rating",
       y = "Density", 
       col = "Genre")
```

The densities are very similar, although dramas appear to be rated a tiny bit higher.


# Question 5

Plotting histograms:
```{r}
ggplot(comedies_dramas,
       aes(av_rating)) +
  geom_histogram() +
  theme_minimal() +
  labs(title = "Distribution of average ratings by genre",
       x = "Average rating",
       y = "Number of shows") +
  facet_wrap(~ genres)
```

This shows that a) there were far more comedies (by our categorization) in the dataset and b) the distribution of low-rated outliers is quite similar.

```{r}
ggplot(comedies_dramas,
       aes(av_rating, color = genres)) +
  geom_freqpoly() +
  theme_minimal() +
  labs(title = "Distribution of average ratings by genre",
       x = "Average rating",
       y = "Number of shows",
       col = "Genres")
```

For cases like this, I'd definitely prefer the frequency plot we plotted last. 


# Question 6

```{r}
ggplot(comedies_dramas,
       aes(av_rating, share)) +
  geom_bin_2d() +
  theme_minimal() +
  labs(title = "Viewership by average rating",
       x = "Average rating for season",
       y = "Share of viewership")
```

This creates a "heat map" which can be a bit more intuitive than a scatterplot when it comes to showing where a distribution is particularly dense. 

```{r}
ggplot(comedies_dramas,
       aes(av_rating, share,
           fill = genres)) +
  geom_bin_2d() +
  theme_minimal() +
  labs(title = "Viewership by average rating",
       x = "Average rating for season",
       y = "Share of viewership",
       fill = "Genre")
```

It looks as if decently-rated comedies had a much higher viewership than corresponding dramas. 

```{r}
comedies_dramas$title[which.max(comedies_dramas$share)]
```

That was apparently "Dekalog". Weird that I never heard of it, looks really interesting though. 






