---
title: "MD 1 Homework"
author: "Christoph Rosa"
date: "2022-09-01"
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---
# Chapter 1

## Question 1:

Installing packages. 

```{r, eval = FALSE}
install.packages("causact")
install.packages("dplyr")
install.packages("igraph")

```

## Question 2

Loading packages:

```{r}
library(causact)
library(dplyr)
library(igraph)
```

Let's check out if calling as_data_frame without specifying a package works on a vector.

```{r}
#df <- as_data_frame(x = c(1,2,3))
```

Shocking. Now for dplyr's as_data_frame:

```{r}
df <- dplyr::as_data_frame(x = c(1,2,3))
glimpse(df)
```

Using dplyr' n_distinct function:

```{r}
x <- c(5,6,2,7,9,1)
dplyr::n_distinct(x)
```

R by default used the function from igraph because it "masked" the dplyr function, simply because igraph was installed later. Had we installed igraph first, R should by default use the dplyr function, as the one from igraph would be "masked" by it. 

## Question 3

```{r}
?n_distinct
```

It calculates the number of distinct values (numbers, strings, etc.) in a vector. 

## Question 4

```{r}
glimpse(baseballData)
```

The dataset has 12,145 rows and five columns. The variables in the column "Home" are factors, the one in "Homescore" are integers. 

## Question 5

```{r}
baseballData[1,]
```

Each row represents one baseball game.

```{r}
baseballData[,2:3] %>% head()
```

The home team and the visiting team in each of the games.

## Question 6

The hockey data:

```{r}
name <-
  c(
    "Wayne Gretzky",
    "Gordie Howe",
    "Jaromir Jagr",
    "Brett Hull",
    "Marcel Dionne",
    "Phil Esposito" ,
    "Mike Gartner",
    "Alex Ovechkin",
    "Mark Messier" ,
    "Steve Yzerman"
  )

goals <- c(894, 801, 766, 741, 731, 717, 708, 700, 694, 692)

year_started <- c(1979, 1946, 1990, 1986, 1971, 1963, 1979, 2005, 1979, 1983)

```

Building the dataframe:

```{r}
df <- tibble( 
  "Name" = name, 
  "Goals" = goals, 
  "Year started" = year_started) 

glimpse(df)
```






