---
title: "MD_4_homework_CR"
author: "Christoph Rosa"
date: "2022-09-09"
output: html_document
---

# Set-up

Installing tidyverse and cleaning up:
```{r}
library(tidyverse)
rm(list = ls())
```


# Question 1

Downloading the NFL data:
```{r}
nflraw <- read_csv("https://raw.githubusercontent.com/NicolasRestrep/223_course/main/Data/nfl_salaries.csv")
glimpse(nflraw)
```


# Question 2

Tidying up our dataset:
```{r}
nfl <- nflraw %>% 
  pivot_longer(names_to = "position",
               values_to = "salary",
               cols = -year)
glimpse(nfl)
```


# Question 3

Making a history of Quarterback salaries:
```{r}
qbs <- nfl %>% filter(position == "Quarterback") 

options(scipen=10000)

ggplot(qbs, aes(salary)) +
  geom_histogram() + 
  theme_minimal() +
  labs(title = "Salaries of NFL quarterbacks",
         x = "",
         y = "Salary in USD")

```

And now, to compare years:
```{r}
ggplot(qbs, aes(salary)) +
  geom_histogram() + 
  theme_minimal() +
  labs(title = "Salaries of NFL quarterbacks",
         x = "",
         y = "Salary in USD") +
  theme(axis.text.x = element_text(angle = 90)) +
  facet_wrap(~ year, nrow = 2)
```

It looks as if over the years, the salaries of the highest paid Quarterbacks have become even higher. It also seems like the very few best-paid players have set themselves apart even from this group of high-earning outliers. 


# Question 4

Calculating average salaries per year and position:
```{r}
avg <- nfl %>% group_by(position, year) %>% 
  summarize(mean = mean(salary, na.rm = TRUE))
glimpse(avg)
```


# Question 5

Let's plot how the salaries of the positions have evolved over the years:

```{r}
ggplot(avg,
       aes(year, mean, 
           group = position,
           color = position)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Average salary by player position",
       x = "Year",
       y = "Average salary",
       col = "Position")
```

With the notable exception of the poor RBs, the average salary for every position has increased more or less steadily over the years.

This increase was steeper for the - on average - better-paid positions (the best-paid positions have been quite consistently Linemen, both on Offense and Defense). This has increased the disparity of the average salaries over time. 




