---
title: "SR_3_homework_CR"
author: "Christoph Rosa"
date: "2022-09-27"
output: html_document
---

# Easy 

## Loading packages and data

First, packages:
```{r}
library(tidyverse)
library(rethinking)
```

And the data given by the book:
```{r}
p_grid <- seq(from=0, to=1, length.out=1000)
prior <- rep( 1, 1000)
likelihood <- dbinom(6, size=9, prob=p_grid)
posterior <- likelihood * prior
posterior <- posterior/sum(posterior) 

set.seed(100)
samples <- sample(p_grid, prob=posterior, size=1e4, replace=TRUE) 
```

## Easy 1

```{r}
options(scipen = 999) # stop R from using scientific notation
sum(samples < 0.2)/length(samples)
```

Only 0.04% of the posterior probability lies below p = 0.2. 


## Easy 2
```{r}
sum(samples > 0.8)/length(samples)
```

About 11.2% of the posterior probability lies above p = 0.8.


## Easy 3

```{r}
sum(samples > 0.2 & samples < 0.8)/length(samples)
```

88.8% of the posterior probability lies between p = 0.2 and p = 0.8. 


## Easy 4

```{r}
quantile(samples, 0.2)
```

20% of the posterior probability lies below p = 0.52. 


## Easy 5

```{r}
quantile(samples, 0.8)
```

20% of the posterior probability lies above p = 0.76. 


## Easy 6

```{r}
rethinking::HPDI(samples, prob = .66)
```

The values p = 0.51 and p = 0.77 contain the narrowest interval equal to 66% of the posterior probability.


## Easy 7

```{r}
PI(samples, prob = 0.66)
```

The 66% percentile compatibility interval lies between p = 0.50 and p = 0.77. 


# Medium 

## Medium 1

Calculating grid approximation analogous to Chapter 2:
```{r}
grid <- seq(0, 1, length.out = 1000)
prior <- rep(1, 1000)

d1 <- tibble(grid, prior) %>% 
  mutate(likelihood = dbinom(8, 15, grid)) %>% 
  mutate(unpost = likelihood * prior) %>% 
  mutate(post = unpost / sum(unpost))

d1 %>% ggplot(aes(x = grid, y = post)) +
  geom_point() +
  geom_line() +
  labs(title = "M1: Grid approximation",
       x = "Probability of water",
       y = "Posterior probability") +
  theme_minimal()
```


## Medium 2

Now, we draw 10k samples:
```{r}
set.seed(100)
samples2 <- d1 %>% slice_sample(n = 1e4, weight_by = post, replace = TRUE) 

rethinking::HPDI(samples2$grid, prob = .9)
```

The narrowest interval equal to 90% of the posterior probability lies between p = 0.33 and p = 0.72. 


## Medium 3

Let's generate 10k simulations of 15 globe tosses:
```{r}
w <- rbinom(1e4, size = 15, prob = samples2$grid)

sum(w == 8)/length(w)
```

The probability of observing 8 waters in 15 tosses is 14.6%. 


## Medium 4

```{r}
w2 <- rbinom(1e4, size = 9, prob = samples2$grid)
sum(w2 == 6)/length(w2)
```

The probability of observing 6 waters in 9 tosses is 17.6%.


## Medium 5

First, our new grid approximation:
```{r}
prior2 <- if_else(grid < .5, 0, 1)

d2 <- tibble(grid, prior2) %>% 
  mutate(likelihood = dbinom(8, 15, grid)) %>% 
  mutate(unpost = likelihood * prior2) %>% 
  mutate(post = unpost / sum(unpost))

d2 %>% ggplot(aes(x = grid, y = post)) +
  geom_point() +
  geom_line() +
  labs(title = "M5: Grid approximation with updated prior",
       x = "Probability of water",
       y = "Posterior probability") +
  theme_minimal()
```

Next, the 90% HDPI: 
```{r}
samples3 <- d2 %>% slice_sample(n = 1e4, weight_by = post, replace = TRUE) 

rethinking::HPDI(samples3$grid, prob = .3)
```

Here, the HDPI lies between p = 0.51 and p = 0.57 - a much narrower interval compared to when we used a flat prior.

```{r}
w3 <- rbinom(1e4, size = 15, prob = samples3$grid)
sum(w3 == 8)/length(w3)

w4 <- rbinom(1e4, size = 9, prob = samples3$grid)
sum(w4 == 6)/length(w4)
```

The updated probability for 8 waters out of 15 tosses is 15.8% (instead of 14.6%), for 6 waters out of 9 tosses it is 23.2 (instead of 17.6%). Both probabilities increased significantly, as the new prior lowered the probability of getting few "waters". 

Let's compare this to the true value of p = 0.7:
```{r}
w5 <- rbinom(1e4, size = 15, prob = 0.7)
sum(w5 == 8)/length(w5)

w6 <- rbinom(1e4, size = 9, prob = 0.7)
sum(w6 == 6)/length(w6)
```

Interestingly, the new prior made our predictions of 8/15 waters even worse, although the flat prior wasn't much better - the actual probability is 8.3%.
On the contrary, the predictions for 6/9 waters became better with the updated prior, closer to the "true" value of 26.9% (but still quite far off).


## Medium 6

I'm unfamiliar with loops and can't think of a better/more efficient way to solve this problem then through simple trial-and-error, so here goes: 
```{r}
tosses = 2200
waters = tosses*0.7

d3 <- tibble(grid, prior) %>% 
  mutate(likelihood = dbinom(waters, tosses, grid)) %>% 
  mutate(unpost = likelihood * prior) %>% 
  mutate(post = unpost / sum(unpost))

samples <- d3 %>% slice_sample(n = 1e4, weight_by = post, replace = TRUE) 

pi <- PI(samples$grid, prob = .99)
pi 
pi[2]-pi[1]
```

One would need to toss the globe approximately a hand-breaking 2,200 times to get an estimate for p with a 99% percentile interval that's only 0.05 wide. 









